---
output:
  html_document:
    code_folding: hide
    number_sections: yes
    theme: cosmo
    toc: yes
    toc_float: yes
---
<style type="text/css">
body .main-container {
  margin-left: 1%;
  margin-right: 1%;
  max-width: 1350px;
}
.jumbotron {
    background-color: #86b0e0!important;
    
}
</style>

<div class="jumbotron">
  <h1>Classifying Brand Preference</h1>
  <p>Rhys Hewer</p>
</div>

# **Executive Summary**

REMEMBER!! Combine the data sets at the end!

## Objective
## Method
## Findings
## Recommendations

# **Initial Data Processing**
```{r message=FALSE}
#load libraries
library(readxl)
library(dplyr)
library(ggplot2)
library(plotly)
library(RColorBrewer)
library(corrplot)
library(caret)
library(tidyr)
library(kableExtra)
library(parallel)
library(doParallel)
library(rattle)
library(rpart)
```
```{r}
#read in data
setwd("C:/Users/rhysh/Google Drive/Data Science/Ubiqum/Project 2/Task 2")
incomplete <- read.csv("SurveyIncomplete.csv")
origdata <- read_xlsx("Survey_Key_and_Complete_Responses_excel.xlsx", sheet = 2)
data <- origdata
```

As we are working across 2 spreadsheets it is important to check to see if they are structured alike or whether manipulation will be required to ensure they have the same features.


```{r}
#Check structure of data to see if alike
names(incomplete) == names(data)
```

The features match between the spreadsheets so no data manipulation is needed in that respect.

```{r}
incomplete %>% sample_n(5)
data %>% sample_n(5) 
```

A quick look at a sample from both spreadsheets shows that they are very similar in composition.

```{r}
str(incomplete)
str(data)
```

Looking at the structure shows that a few changes are needed in respect of the data types.

Overall, however, the datasets are sufficiently similar. I will take the strategy of splitting into training, test and final prediction sets. 
The training and testing sets will come from the complete responses data. Final prediction from the incomplete responses.

The exploratory data analysis will be performed on the complete data but any data transformations taking place on the training/testing sets will also need to be made on the Final prediction data prior to modelling this.


```{r}
#check for NAs
data %>% is.na() %>% sum()
incomplete %>% is.na() %>% sum()
```

There are no missing values in either data set.

```{r}
#data types
data$elevel <- data$elevel %>% as.factor()
data$car <- data$car %>% as.factor()
data$zipcode <- data$zipcode %>% as.factor()
data$brand <- data$brand %>% as.factor()
```

Education level, car owned, zip code and brand preference are converted from numeric to factors.

```{r}
##check for outliers
numericVars <- Filter(is.numeric, data)
outliers <- numericVars %>% sapply(function(x) boxplot(x, plot=FALSE)$out) %>% str()
```

There are no outliers.

# **Exploratory Data Analysis**

##Feature Histograms {.tabset}

### Brand Preference Histogram
```{r}
#EDA

##Key feature is brand preference, begin with exploring this value.

g6 <- ggplot(data, aes(brand, fill = brand)) +
        geom_bar() +
        theme_bw() +
        scale_fill_brewer(palette="Dark2") +
        xlab("Brand Preference") + 
        ylab("Frequency") + 
        ggtitle("Brand Preference Frequencies")
g6
```

* We see a clear preference for brand 1 over brand 0. 

### Histogram of other variables
```{r}
##review other histograms for skewdness
histData <- origdata %>% select(-brand)
g8 <- ggplot(gather(histData), aes(value)) + 
        geom_histogram(bins = 10, fill = "#D95F02", colour = "white") + 
        theme_bw() +
        facet_wrap(~key, scales = 'free_x') +
        xlab("Value") + 
        ylab("Count") + 
        ggtitle("Histograms of Numeric Variables")
g8
```

* Across the remaining features we see no extreme skewing or noteworthy patterns.

## Brand Choice Plotting {.tabset}

### Brand Choice v Salary
```{r}
## Plot Brand choice v other variables
g1 <- ggplot(data, aes(brand, salary, fill = brand)) +
        geom_violin() +
        theme_bw() +
        scale_fill_brewer(palette="Dark2") +
        xlab("Brand Preference") + 
        ylab("Salary ($)") + 
        ggtitle("Brand Preference v Salary")
g1 <- ggplotly(g1)
g1
```

* There is a clear pattern: salaries ranging between 45k - 100k seem to have a preference for brand 0. Salaries outside of this range seem to have a preference for brand 1.

### Brand Choice v Age
```{r}
g2 <- ggplot(data, aes(brand, age, fill = brand)) +
        geom_violin() +
        theme_bw() +
        scale_fill_brewer(palette="Dark2") +
        xlab("Brand Preference") + 
        ylab("Age") + 
        ggtitle("Brand Preference v Age")
g2 <- ggplotly(g2)
g2
```

* There is no noteworthy pattern between brand preference and age.

### Brand Choice v Education Level
```{r}
g3 <- ggplot(data, aes(brand, elevel, colour = brand)) +
        geom_count() +
        theme_bw() +
        scale_color_brewer(palette="Dark2") +
        xlab("Brand Preference") + 
        ylab("Education Level") + 
        ggtitle("Brand Preference v Education Level")
g3
```

* The general preference for brand 1 is shown in this plot but there also seems to be a consistent preference for brand 1 regardless of education level whereas there seems to be more variation within preference for brand 0 based on the education level.

### Brand Choice v Car 
```{r}
g4 <- ggplot(data, aes(brand, car, colour = brand)) +
        geom_count() +
        theme_bw() +
        scale_color_brewer(palette="Dark2") +
        xlab("Brand Preference") + 
        ylab("Car") + 
        ggtitle("Brand Preference v Car")
g4
```

* The general preference for brand 1 is shown in this plot but there also seems to be a consistent preference for brand 1 regardless of car whereas there seems to be more variation within preference for brand 0 based on the car owned.

### Brand Choice v Zip Code
```{r}
g5 <- ggplot(data, aes(brand, zipcode, colour = brand)) +
        geom_count() +
        theme_bw() +
        scale_color_brewer(palette="Dark2") +
        xlab("Brand Preference") + 
        ylab("Zip Code") + 
        ggtitle("Brand Preference v Zip Code")
g5
```

* The general preference for brand 1 is shown in this plot but there also seems to be a consistent preference for brand 1 regardless of zip code whereas there seems to be more variation within preference for brand 0 based on the car owned.

### Brand Choice v Credit
```{r}
g7 <- ggplot(data, aes(brand, credit, fill = brand)) +
        geom_violin() +
        theme_bw() +
        scale_fill_brewer(palette="Dark2") +
        xlab("Brand Preference") + 
        ylab("Credit") + 
        ggtitle("Brand Preference v Credit")
g7 <- ggplotly(g7)
g7
```

* There is no noteworthy pattern between brand preference and credit.

## Initial Hypothesis

There is a general preference for brand 1 and salary seems the key feature.

Age and credit seem to have very limited influence on brand preference. Whilst there are patterns of preference within education level, car and zip code, salary appears to have the most striking impact.

Salaries ranging between 45k - 100k seem to have a preference for brand 0. Salaries outside of this range seem to have a preference for brand 1.


## Feature Selection

### Decision Tree Variable Importance

We can review the initial hypothesis and provide some information on which to make feature selection decisions using a decision tree.

```{r}
featureDT <- rpart(brand ~ ., data = data)
featureDT$variable.importance
fancyRpartPlot(featureDT)
```

The decision tree is formed exclusively of age and salary, suggesting these are the key features. This is reinforced by the variable importance information from within the model. 

### Supplementary Exploratory Data Analysis

```{r}
g12 <- ggplot(data, aes(salary, age, colour = brand)) +
        geom_point(show.legend = FALSE) +
        facet_grid(brand ~ .) +
        theme_bw() +
        scale_color_brewer(palette="Dark2") +
        xlab("Salary") + 
        ylab("Age") + 
        ggtitle("Salary v Age by Brand Preference")
g12
```

We see that the interaction of age and salary do play a role in brand preference. There are clearly defined blocks of preference depending on age and salary. For example, between ages 60-80 earning between 25k-80k have a distinct preference for brand 0.

I have updated the initial hypothesis to now consider that age and salary are the key features in determining brand preference.

### Colinearity and Variance
```{r}
##Review correlation matrix
corrMatrix <- origdata %>% cor()
corrMatrix %>% corrplot.mixed()
```

Reviewing the correlation plot we see very little correlation between the features. This means that colinearity is not an issue that needs to be addressed.

```{r}
#near zero variance
nzv <- data %>% nearZeroVar(saveMetrics = TRUE)
nzv
```

There are no features with near zero variance.

### Feature Selection Conclusion

The Random Forest algorithm has built-in feature selection so no feature selection required from that perspective.

For KNN I will take both a fully-featured dataset and a concentrated dataset (brand, salary, age) into the modelling phase and compare results.
 
# **Modelling**

## Modelling Preparation

```{r}
#Creating Testing/Training sets
set.seed(111)
trainIndex <- createDataPartition(iris$Species, p = 0.75, list = FALSE)

training <- data[ trainIndex,]
testing  <- data[-trainIndex,]

concData <- data %>% select(brand, age, salary)
training.conc <- concData[ trainIndex,]
testing.conc <- concData[-trainIndex,]

#set up parallel processing (requires parallel and doParallel libraries)
cluster <- makeCluster(detectCores() - 1) 
registerDoParallel(cluster)

#Cross Validation 10 fold
fitControl<- trainControl(method = "cv", number = 10, savePredictions = TRUE, allowParallel = TRUE)
```

A test and traing set are created with all features. An additional training set is created with just Brand, Age and Salary. 10-fold cross validation is being used and parallel processing is being leveraged.

## Modelling K-Nearest Neighbour 

### Training KNN {.tabset}

#### KNN - All Features
```{r}
# Deploy KNN model 
model.KNN.brand <- train(brand ~ ., data = training, 
                         method = "knn", trControl = fitControl)
model.KNN.brand
```

#### KNN - Concentrated Features
```{r}
# Deploy KNN model with concentrated feature set
model.KNN.brand.conc <- train(brand ~ ., data = training.conc, 
                              method = "knn", trControl = fitControl)
model.KNN.brand.conc
```

#### KNN - All Features, Scaled and Centred
```{r}
# Deploy KNN model with scaled features
model.KNN.brand.scale <- train(brand ~ ., data = training, 
                         method = "knn", trControl = fitControl, 
                         preProcess = c("center", "scale"))
model.KNN.brand.scale
```


#### KNN - Concentrated Features, Scaled and Centred
```{r}
# Deploy KNN model with scaled features and concentrated feature set
model.KNN.brand.conc.scale <- train(brand ~ ., data = training.conc, 
                              method = "knn", trControl = fitControl, 
                              preProcess = c("center", "scale"))
model.KNN.brand.conc.scale
```

#### KNN - best model tuned

```{r}
#tune best KNN model further
tGridKNN <- expand.grid(k = c(1,2,3,4,5,6))
finalKNN <- train(brand ~ ., data = training.conc,
                  method = "knn", trControl = fitControl,
                  preProcess = c("center", "scale"),
                  tuneGrid = tGridKNN)
finalKNN
```

By tuning the model I was able to increase the accuracy on the training sent to 91% and the Kappa to 0.78. This brings concerns of overfitting. 


### Testing KNN

```{r}
#Predictions on the test set (model = model name, testing = test set)
predictions.KNN <- predict(finalKNN, testing)
testing$predictions.KNN <- predictions.KNN


#Confusion matrix
confMatrixKNN <- confusionMatrix(testing$brand, testing$predictions.KNN)
confMatrixKNN
fourfoldplot(confMatrix$table, conf.level = 0, margin = 1, main = "Confusion Matrix KNN")
```

The Accuracy of the model on the test set was 87% with a kappa of 0.72. These seem to be satisfactory outcomes. Reviewing the confusion matrix we see that the model is doing a fairly good job of classifying brand preference with a slight tendency to overpredict brand 0.

## Modelling Random Forest

### Training Random Forest {.tabset}

#### Random Forest on full feature set
```{r}
# Deploy RF model 
tGridRF <- expand.grid(mtry = c(24,30,35))
model.RF.brand.tune <- train(brand ~ ., data = training, 
                        method = "rf", trControl = fitControl,
                        tuneGrid = tGridRF)
model.RF.brand.tune
```

#### Random Forest on concentrated feature set
```{r}
## Deploy RF model on concentrated feature set
model.RF.brand.conc <- train(brand ~ ., data = training.conc, 
                         method = "rf", trControl = fitControl)
model.RF.brand.conc 
varImp(model.RF.brand.conc)

```

### Testing Random Forest

```{r}
## Test Random Forest
predictions.RF <- predict(model.RF.brand.conc, testing)
testing$predictions.RF <- predictions.RF

#Confusion matrix
confMatrixRF <- confusionMatrix(testing$brand, testing$predictions.RF)
confMatrixRF
fourfoldplot(confMatrixRF$table, conf.level = 0, margin = 1, main = "Confusion Matrix RF")
```

## Modelling Conclusions

# **Conclusions**

